\HeaderA{basehaz.gbm}{Baseline hazard function}{basehaz.gbm}
\keyword{methods}{basehaz.gbm}
\keyword{survival}{basehaz.gbm}
\begin{Description}\relax
Computes the Breslow estimator of the baseline hazard function for a proportional hazard regression model
\end{Description}
\begin{Usage}
\begin{verbatim}
basehaz.gbm(t, delta, f.x, 
            t.eval = NULL, 
            smooth = FALSE, 
            cumulative = TRUE)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{t}] the survival times 
\item[\code{delta}] the censoring indicator 
\item[\code{f.x}] the predicted values of the regression model on the log hazard scale 
\item[\code{t.eval}] values at which the baseline hazard will be evaluated 
\item[\code{smooth}] if \code{TRUE} \code{basehaz.gbm} will smooth the estimated baseline hazard using Friedman's super smoother \code{\LinkA{supsmu}{supsmu}}
\item[\code{cumulative}] if \code{TRUE} the cumulative survival function will be computed 
\end{ldescription}
\end{Arguments}
\begin{Details}\relax
The proportional hazard model assumes h(t|x)=lambda(t)*exp(f(x)). \code{\LinkA{gbm}{gbm}} can estimate the f(x) component via partial likelihood. After estimating f(x), \code{basehaz.gbm} can compute the a nonparametric estimate of lambda(t).
\end{Details}
\begin{Value}
a vector of length equal to the length of t (or of length \code{t.eval} if \code{t.eval} is not \code{NULL} containing the baseline hazard evaluated at t (or at \code{t.eval} if \code{t.eval} is not \code{NULL}). If \code{cumulative} is set to \code{TRUE} then the returned vector evaluates the cumulative hazard function at those values.
\end{Value}
\begin{Author}\relax
Greg Ridgeway \email{gregr@rand.org}
\end{Author}
\begin{References}\relax
N. Breslow (1972). "Disussion of `Regression Models and Life-Tables' by D.R. Cox," Journal of the Royal Statistical Society, Series B, 34(2):216-217.

N. Breslow (1974). "Covariance analysis of censored survival data," Biometrics 30:89-99.
\end{References}
\begin{SeeAlso}\relax
\code{\LinkA{survfit}{survfit}}, \code{\LinkA{gbm}{gbm}}
\end{SeeAlso}

